Malacology: A Programmable Storage System

### 2017-02-13 paper/paper.tex
paper: version 0; addresses reviewer confusion - clarify what was implemented - describe what makes a programmable storage system - organize motivation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment for diff
\includecomment{notes}
\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for camera-ready
%\excludecomment{notes}
%\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
%\newcommand{\oldcomment}[1]{}
%\newcommand{\addressesissue}[1]{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{Malacology: A Programmable Storage System}
%\authorinfo{Paper ID: 165}          {Total Pages: 12}
\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, Carlos Maltzahn, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre}
           {\{msevilla, jayhawk, ivo, cmaltzahn\}@soe.ucsc.edu, \{palvaro, finkelstein, jlefevre\}@ucsc.edu}
\date{}
\begin{document}

@@

\begin{notes}
\textcolor{red}{
\noindent Issues raised by reviewers:
\begin{enumerate}
  \item clarify what was implemented: diagram of interfaces and prose
  specifying the difference between Malacology and prior work (reviewers B, C,
  D, E)
  \item describe what makes a programmable storage system (reviewers C, D)
  \item organize motivation: group interfaces into categories (e.g., core
  functionality, features, and performance optimizations) and describe other
  applications we can build with Malacology (reviewers B, C)
  \item clarify terminology: service vs. file system metadata, storage vs.
  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
  B, C)
  \item address security, access control, and the safety; specify how it
  affects composability, performance, and multi-tenancy (reviewer D)
\end{enumerate}
}
\end{notes}

@@

The main challenge of designing a programmable storage system is choosing the
right internal abstractions and picking the correct layers for exposing them.
\oldcomment{In this paper, we do not present an exhaustive list of possible
internal abstractions nor do we contend that the abstractions we choose provide
the best trade-offs for all applications.} 
\newcomment{A programmable storage is not defined by what abstractions are
exposed, rather a programmable storage system adheres to the design approach of
choosing interfaces so administrators can have better control of the storage
system. The interfaces presented in this paper are abstractions that we found
useful for building ZLog and Mantle, yet they may not provide the best
trade-offs for all higher-level services.}
\addressesissue{2}
For example, if consensus is correctly exposed one could implement
high-level features like versioning, serialization, or various flavors of
strongly consistent data management on top; but perhaps a low-level consensus
interface is suited well for a particular set of applications.  These questions
are not answered in this paper and instead we focus on showing the feasibility
of building such a system, given advances in the quality and robustness of
today's storage stacks.

@@

\begin{table*}
\centering\small
\begin{tabular}{ l | c | l | l | l }
%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Interface}                      &
\textbf{Sec}                            &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
Service Metadata
  & \S\ref{sec:service-metadata-interface}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
  & cluster state management~\cite{website:ceph-mon}
  & consensus/consistency
  \\
Data I/O
  & \S\ref{sec:data-io-interface}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction/atomicity
  \\
Shared Resource
  & \S\ref{sec:distributed-metadata-interfaces}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization/batching
  \\
File Type
  & \S\ref{sec:distributed-metadata-interfaces}
  & ?????
  & file striping strategy
  & data/metadata access
  \\
Load Balancing
  & \S\ref{sec:distributed-metadata-interfaces}
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration/sampling
  \\
Durability
  & \S\ref{sec:durability}
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & persistence/safety
  \\
\end{tabular}
\caption{Common internal abstractions. The same ``internal abstractions" are
common across large-scale systems because they provide primitives that solve
general distributed systems problems.  Here we list examples of what these
internal abstractions are used for in ``production systems" and in Ceph.
Malacology provides these internal abstractions as interfaces
\newcomment{(Section~\ref{sec:malacology})}
that higher level
\oldcomment{applications}
\newcomment{services (Section~\ref{sec:services})}
can use.
\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
paper.}}
\addressesissue{1}
\label{table:examples}
\end{table*}

@@

\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
metadata server) data structures called ``maps'' that record membership and
status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as Paxos are a common technique for maintaining consistent
data versions, and are a useful system to expose.

@@

\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
time-varying service metadata as a service rather than a hidden internal
component. Malacology provides a generic API for adding arbitrary values to
existing subsystem cluster maps. As a consequence of this, applications can
define simple but useful service-specific logic to the strongly-consistent
interface, such as authorization control (just specific clients can write new
values) or triggering actions based on specific values (e.g. sanitize values).
The higher level services we implement in \S\ref{sec:services} make use of this
functionality to register, version and propagate dynamic code (Lua scripts) for
new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
policies in the load balancer \S\ref{malacology:mds}.  Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughout the cluster so that
clients are properly synchronized with the latest interfaces.

\newcomment{\\ \noindent \it{\textbf{Impact: }} provides core functionality
because it lets daemons come to consensus on system critical state. Bugs in the
internal subsystems or ommitting this from services that need this type of
consistency affects correctness.}
\addressesissue{3, 5}

@@

\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
An object interface is a plugin structured similarly to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and writing to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
interfaces may be transactionally composed along with application
specific logic to create semantically rich interfaces. An example would be
an interface that atomically updates a matrix stored in the bytestream and an
index of the matrix stored in the key-value database.\\

@@

\oldcomment{\noindent Shared resource interface.}
\subsubsection{Shared Resource Interface}

@@

\newcomment{\\ \noindent\it{\textbf{Ceph Implementation: }}}
In Ceph, the locking service implements a capability-based system that
expresses what data and metadata clients are allowed to access as well as what
state they may cache and modify locally.  While designed for the file
abstraction, indexing, locking, and caching are all common services that are
useful to a broad spectrum of applications.  Distributed applications that
share centralized resources (e.g. a database or directory) face similar
challenges which are often solved using application-specific sharding.

@@

\oldcomment{\noindent File type interface.}
\subsubsection{File Type Interface}
The metadata service manages a POSIX file system hierarchy where files and
directories are represented as inode data structures and expose a POSIX file
interface. Applications that manage large amounts of metadata (e.g. users or
database snapshots) often require a naming service. We allow new inode types to
be defined such that applications can create domain-specific interfaces to
inodes that may modify locking and capability policies. We will show how this
is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
on Malacology.

@@

\oldcomment{\noindent \bf Load balancing interface.} 
\subsubsection{Load Balancing Interface}

Ceph also addresses the challenge of balancing metadata load with a separate
metadata cluster. This cluster uses load balancing policies to migrate
directory inodes around the cluster to alleviate overload at single
nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
state (e.g.  CPU and memory utilization) and statistics collected by the
cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
partitioning to move variable sized namespace subtrees. These units can be
shipped anywhere (i.e., to any metadata server of any capacity) at any time for
any reason. The original balancer was designed with hard-coded policies and
tunables.\\

@@

Ceph provides storage by striping and replicating data across
RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
protects data using common techniques such as erasure
coding, replication, and scrubbing.  Furthermore, many of these features
are implemented using a peer-to-peer protocol that allows object storage
devices to operate autonomously without a centralized coordinator.  For
example, when the number of placement groups change, the OSDs re-balance and
re-shard data in the background in a process called placement group splitting
in which OSDs communicate directly with each other to cover a new state.
In order to reduce load on the monitoring service, Ceph OSDs use a gossip
protocol to efficiently propagate changes to cluster maps throughout the
system, and autonomously initiate recovery mechanisms when failures are
discovered.\\

@@

\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
adds data safety and persistence to system metadata; while nice to have it does
effect correctness.}
\addressesissue{3}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment for diff
%\includecomment{notes}
%\newcommand{\newcomment}[1]{{\textcolor{red}{\textbf{#1}}}}
%\newcommand{\oldcomment}[1]{{\textcolor{blue}{\textbf{\sout{#1}}}}}
%\newcommand{\addressesissue}[1]{{\textcolor{red}{{\it (Addresses issue(s): {#1})}}}}

% Uncomment for camera-ready
\excludecomment{notes}
\newcommand{\newcomment}[1]{{\textcolor{black}{#1}}}
\newcommand{\oldcomment}[1]{}
\newcommand{\addressesissue}[1]{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{Malacology: A Programmable Storage System}
%\authorinfo{Paper ID: 165}          {Total Pages: 12}
\authorinfo{Michael A. Sevilla, Noah Watkins, Ivo Jimenez, \\Peter Alvaro, Shel Finkelstein, Jeff LeFevre, Carlos Maltzahn}
           {\{msevilla, jayhawk, ivo\}@soe.ucsc.edu, \{palvaro, shel, jlefevre, carlosm\}@ucsc.edu}
\date{}
\begin{document}

@@

\begin{notes}
\textcolor{red}{
\noindent Issues raised by reviewers, paper augmented as noted inline (red/blue text):
\begin{enumerate}
  \item clarify what was implemented: diagram of interfaces and prose
  specifying the difference between Malacology and prior work (reviewers B, C,
  D, E)
  \item describe what makes a programmable storage system (reviewers C, D)
  \item organize motivation: group interfaces into categories (e.g., core
  functionality, features, and performance optimizations) and describe other
  applications we can build with Malacology (reviewers B, C)
  \item clarify terminology: service vs. file system metadata, storage vs.
  file system, dirty vs. clean slate approaches, soft-state caching (reviewers
  B, C)
  \item address security, access control, and the safety; specify how it
  affects composability, performance, and multi-tenancy (reviewer D)
\end{enumerate}
}
\end{notes}

@@

The main challenge of designing a programmable storage system is choosing the
right internal abstractions and picking the correct layers for exposing them.
\oldcomment{In this paper, we do not present an exhaustive list of possible
internal abstractions nor do we contend that the abstractions we choose provide
the best trade-offs for all applications.} 
\newcomment{A programmable storage is not defined by what abstractions are
exposed, rather a programmable storage system adheres to the design approach of
choosing interfaces so administrators can have better control of the storage
system. The interfaces presented in this paper are abstractions that we found
useful for building our prototype services ZLog and Mantle, yet they may not provide the best
trade-offs for all higher-level services.}
\addressesissue{2}
For example, if consensus is correctly exposed one could implement
high-level features like versioning, serialization, or various flavors of
strongly consistent data management on top; but perhaps a low-level consensus
interface is suited well for a particular set of applications.  These questions
are not answered in this paper and instead we focus on showing the feasibility
of building such a system, given advances in the quality and robustness of
today's storage stacks.

@@

\begin{table*}
\centering\small
\begin{tabular}{ l | c | l | l | l }
%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Interface}                      &
\textbf{Section}                        &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
Service Metadata
  & \S\ref{sec:service-metadata-interface}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
  & cluster state management~\cite{website:ceph-mon}
  & consensus/consistency
  \\
Data I/O
  & \S\ref{sec:data-io-interface}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction/atomicity
  \\
Shared Resource
  & \S\ref{sec:shared-resource-interface}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization/batching
  \\
File Type
  & \S\ref{sec:file-type-interface}
  & ?????
  & file striping strategy
  & data/metadata access
  \\
Load Balancing
  & \S\ref{sec:load-balancing-interface}
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration/sampling
  \\
Durability
  & \S\ref{sec:durability}
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & persistence/safety
  \\
\end{tabular}
\caption{Common internal abstractions. The same ``internal abstractions" are
common across large-scale systems because they provide primitives that solve
general distributed systems problems.  Here we list examples of what these
internal abstractions are used for in ``production systems" and in Ceph.
Malacology provides these internal abstractions as interfaces
\newcomment{(Section~\ref{sec:malacology})}
that higher level
\oldcomment{applications}
\newcomment{services (Section~\ref{sec:services})}
can use.
\oldcomment{; {\it italicized abstractions} are interfaces discussed in this
paper.}}
\addressesissue{1}
\label{table:examples}
\end{table*}

@@

\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
metadata server) data structures called ``maps'' that record membership and
status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as Paxos are a common technique for maintaining consistent
data versions, and are a useful system to expose.

@@

\noindent{\it{\textbf{Malacology: }}} Malacology exposes a strongly-consistent view of
time-varying service metadata as a service rather than a hidden internal
component. Malacology provides a generic API for adding arbitrary values to
existing subsystem cluster maps. As a consequence of this, applications can
define simple but useful service-specific logic to the strongly-consistent
interface, such as authorization control (just specific clients can write new
values) or triggering actions based on specific values (e.g. sanitize values).
The higher level services we implement in \S\ref{sec:services} make use of this
functionality to register, version and propagate dynamic code (Lua scripts) for
new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
policies in the load balancer \S\ref{malacology:mds}.  Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughout the cluster so that
clients are properly synchronized with the latest interfaces.\\

\noindent\newcomment{{\it{\textbf{Impact: }}} provides core functionality
because it lets daemons come to consensus on system critical state. Bugs in the
internal subsystems or ommitting this from services that need this type of
consistency affects correctness.}
\addressesissue{3, 5}

@@

\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
An object interface is a plugin structured similarly to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and writing to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
interfaces may be transactionally composed along with application
specific logic to create semantically rich interfaces. An example would be
an interface that atomically updates a matrix stored in the bytestream and an
index of the matrix stored in the key-value database.\\

@@

\oldcomment{\noindent Shared resource interface.}
\subsubsection{Shared Resource Interface}
\label{sec:shared-resource-interface}

@@

\newcomment{\\ \noindent\it{\textbf{Existing Ceph Implementation: }}}
In Ceph, the locking service implements a capability-based system that
expresses what data and metadata clients are allowed to access as well as what
state they may cache and modify locally.  While designed for the file
abstraction, indexing, locking, and caching are all common services that are
useful to a broad spectrum of applications.  Distributed applications that
share centralized resources (e.g. a database or directory) face similar
challenges which are often solved using application-specific sharding.

@@

\oldcomment{\noindent File type interface.}
\subsubsection{Malacology: File Type Interface}
\label{sec:file-type-interface}

Applications that manage large amounts of metadata (e.g. users or database
snapshots) often require a naming service.  The metadata service manages a
POSIX file system hierarchy where files and directories are represented as
inode data structures and expose a POSIX file interface.

\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}} CephFS
is the POSIX compliant file system that uses Ceph.  Inodes are quite large (1KB
for an inode, 400 bytes for a directory entry, and 700 bytes for a directory)
and contain CephFS-specific policies like how to stripe data across RADOS.}

\newcomment{\\ \noindent {\it{\textbf{Malacology: }}}} We allow new inode types to
be defined such that applications can create domain-specific interfaces to
inodes that may modify locking and capability policies. We will show how this
is used in Section~\ref{sec:seq} when we discuss a distributed shared-log built
on Malacology.

@@

\oldcomment{\noindent \bf Load balancing interface.} 
\subsubsection{Load Balancing Interface}
\label{sec:load-balancing-interface}

\newcomment{Many large scale storage systems separate metadata and data IO so
that the corresponding services can scale independently. Metadata requests
transfer small amounts of data and they happen relatively frequently so many
systems employ separate metadata clusters.}

\newcomment{\\ \noindent {\it{\textbf{Existing Ceph Implementation: }}}} Ceph
also addresses the challenge of balancing metadata load with a separate
metadata cluster. This cluster uses load balancing policies to migrate
directory inodes around the cluster to alleviate overload at single
nodes~\cite{weil:sc2004-dyn-metadata}.  The policies use metrics based on
system state (e.g.  CPU and memory utilization) and statistics collected by the
cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
partitioning to move variable sized namespace subtrees. These units can be
shipped anywhere (i.e., to any metadata server of any capacity) at any time for
any reason. The original balancer was designed with hard-coded policies and
tunables.\\

@@

\newcomment{Object stores protect data using techniques like erasure coding,
replication, and data scrubbing. For scalability, many of these features are
implemented using a peer-to-peer protocol that allows object storage devices to
operate autonomously without a centralized coordinator.\\}

\noindent{\it{\textbf{Existing Ceph Implementation: }}}
Ceph provides storage by striping and replicating data across
RADOS~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
protects data using common techniques such as erasure coding, replication, and
scrubbing.  example, when the number of placement groups change, the OSDs
re-balance and re-shard data in the background in a process called placement
group splitting in which OSDs communicate directly with each other to cover a
new state.  In order to reduce load on the monitoring service, Ceph OSDs use a
gossip protocol to efficiently propagate changes to cluster maps throughout the
system, and autonomously initiate recovery mechanisms when failures are
discovered.\\

@@

\newcomment{\\ \noindent {\it{\textbf{Impact: }}} this is a feature because it
adds data safety and persistence to system metadata; while nice to have it does
not necessarily effect correctness.}
\addressesissue{3}
